Brownian local times

### 2015-10-14 notes/brownian-local-times.Rmd
another approx
.

@@

Also note that for large $|x|$, by the asymptotics for $1-\Phi(x)$,
$$
\begin{aligned}
    \zeta(x,r) = 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) .
\end{aligned}
$$

@@

Here's the cline shape:
```{r show_cline}
pcline <- function (x,r,tau=1,sigma=1) {
    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
}
xx <- seq(-15,15,length.out=500)
for (tau in c(5,50)) {
    plot(xx, pcline(xx,r=1), type='n', 
            xlab="distance from cline center", ylab="prob inherited from the left", 
            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
    rr <- seq(0,0.2,length.out=11)
    pp <- sapply( rr, function (r) { pcline(xx,r=r,tau=tau) } )
    matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
    legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
}
```

@@

Bazykin's approximation to a cline is
$$
p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
$$
Here's what that looks like:
```{r tanh_cline}
tanh_cline <- function (x,s,sigma=1) {
    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
}
plot(xx, tanh_cline(xx,s=0.04) )
```

@@

```{r sim_lineage}
drift <- function (x,s,sigma=1) {
    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
}
nwalks <- 1e4
ntimes <- 1e3
s <- 0.1
# want total time to be 1, so:
eps <- 1/sqrt(ntimes)
rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=(ntimes-1), ncol=nwalks ))
for (k in 2:ntimes) {
    rws[k,] <- rws[k-1,]+rws[k,]+drift(rws[k-1,],s)
}
layout( t(1:2), widths=c(3,1) )
matplot(rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```

>>>>

*Useful functions used below:*
```{r setup_fns}
tanh_cline <- function (x,s,sigma=1) {
    (1/2)*(1+tanh(-2*x*sqrt(s)/sigma))
}
pcline <- function (x,r,tau=1,sigma=1) {
    (1/2)*( 1 - sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau) )
}
```

@@

Also note that by the asymptotics for $1-\Phi(x)$,
$$
\begin{aligned}
    \zeta(x,r) \approx 1 - \sqrt{\frac{2}{\pi}} e^{-x^2/2} \left( \frac{r}{x+r} \right) \qquad \text{for large }|x| .
\end{aligned}
$$

@@

Here's the cline shape:
```{r show_cline}
pcline
xx <- seq(-15,15,length.out=500)
for (tau in c(5,50)) {
    plot(xx, pcline(xx,r=1), type='n', 
            xlab="distance from cline center", ylab="prob inherited from the left", 
            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
    rr <- seq(0,0.2,length.out=11)
    pp <- sapply( rr, function (r) { pcline(xx,r=r,tau=tau) } )
    matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
    legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
}
```

@@

Bazykin's approximation to a cline is
$$
p(x) = \frac{1}{2} (1+\tanh(-2x\sqrt{s}/\sigma))
$$
Here's what that looks like:
```{r tanh_cline}
tanh_cline
plot(xx, tanh_cline(xx,s=0.04) )
```

@@

```{r sim_lineage}
drift <- function (x,s,sigma=1) {
    -(2*sqrt(s)/sigma) * (1-tanh(-2*x*sqrt(s)/sigma)^2) / (1+tanh(-2*x*sqrt(s)/sigma))
}
nwalks <- 1e4
ntimes <- 1e3
s <- 0.1
xx <- seq(-15,15,length.out=500)
# want total time to be 1, so:
eps <- 1/sqrt(ntimes)
rws <- rbind(0,matrix( rnorm(nwalks*(ntimes-1))*eps, nrow=(ntimes-1), ncol=nwalks ))
for (k in 2:ntimes) {
    rws[k,] <- rws[k-1,]+rws[k,]+drift(rws[k-1,],s)
}
layout( t(1:2), widths=c(3,1) )
matplot(rws[,1:80],type='l', ylab='position', xlab='time', ylim=range(rws) )
plot( tanh_cline(xx,s=s), xx, col='green', lwd=2, type='l', ylim=range(rws), ylab='' )
lines( (1-s)*(1-tanh_cline(xx,s=s))/(1-s*(1-tanh_cline(xx,s=s))), xx, lwd=2 )
```
